A 38 Million Year Old Mini-Neptune in the Kepler Field

### 2021-09-07 paper/ms.tex
update star table

The K2 and TESS missions have now enabled the detection of about ten
close-in planets younger than 100 million years, all smaller than
Jupiter
\citep{Mann_K2_33b_2016,David_et_al_2017,david_four_2019,newton_tess_2019,bouma_cluster_2020,plavchan_planet_2020,rizzuto_tess_2020}.
The Kepler mission however has not yielded any planets with precise
ages below one gigayear \citep{Meibom_et_al_2013}.  The reason is that
during the main Kepler mission (2009--2013), only four open
clusters were known in the Kepler field: NGC\,6866, NGC\,6811,
NGC\,6819, and NGC\,6791, with ages spanning 0.7\,Gyr to 9\,Gyr
\citep{meibom_kepler_2011}.  Since that time, analyses of the
kinematic, photometric, and astrometric Gaia data have expanded our
knowledge of open cluster and moving group memberships \citep[{\it
e.g.},][]{cantatgaudin_gaia_2018,zari_3d_2018,kounkel_untangling_2019,Meingast2021,kerr_stars_2021}.
As part of our Cluster Difference Imaging Photometric Survey (CDIPS,
\citealt{bouma_cdipsI_2019}), we concatenated the available analyses
from the literature, which yielded a list of candidate young and
age-dated stars (see Appendix~\ref{app:targetlist}).

@@

Ultimately, the main advance of this work was a precise measurement of
the age of \pn.   This age measurement was enabled by identifying the
connection of the star to the \cn\ using the Gaia data.  Table~\ref{tab:v05}
should enable similar crossmatches to
already known planets, as well as forthcoming planets from the TESS
mission \citep{ricker_transiting_2015,guerrero_tess_2021}.  An
alternative path toward age-dating more planets could be to identify kinematic associations
around known exoplanet host stars using positions and tangential velocities
from Gaia, and to then verify these associations with ancilliary
rotation periods and spectroscopy \citep[{\it e.g.},][]{tofflemire_tess_2021}.

@@

L.G.B{.} and J.L.C{.} are grateful to T{.}~David for help with the
transit fitting, to R{.}~Kerr for kindly providing us with the
\citet{kerr_stars_2021} membership list prior to its publication.
L.G.B{.}, L.H{.}, and N.N{.} are also grateful to K{.}~Collins for
helping resolve the scheduling conflict that would have otherwise
prevented the MuSCAT3 observations.
% is grateful to G.~Zhou, B{.}~Tofflemire, A{.}~McWilliam,
% E{.}~Newton, M{.}~Kounkel, A{.}~Kraus, L{.}~Hillenbrand, and
% K{.}~Hawkins for the discussions on young stars, rotation, and lithium
% that encouraged this analysis.
%
L.G.B{.} acknowledges support by the TESS GI Program, programs
G011103 and G022117, through NASA grants 80NSSC19K0386 and
80NSSC19K1728.
%
L.G.B{.} was also supported by a Charlotte Elizabeth Procter
Fellowship from Princeton University.
%
The archival Keck/NIRC2 imaging was acquired by programs 2015A/N301N2L
(PI: A.~Kraus) and 2019A/N069 (PI: E.~Petigura).
%
In addition, this paper is based in part on observations made with the
MuSCAT3 instrument, developed by the Astrobiology Center and under
financial supports by JSPS KAKENHI (JP18H05439) and JST PRESTO
(JPMJPR1775), at Faulkes Telescope North on Maui, HI, operated by the
Las Cumbres Observatory.
%
This work is partly supported by JSPS KAKENHI Grant Numbers 22000005,
JP15H02063, JP17H04574, JP18H05439, JP18H05442, JST PRESTO Grant
Number JPMJPR1775, the Astrobiology Center of National Institutes of
Natural Sciences (NINS) (Grant Number AB031010).
%
% ACKNOWLEDGE PFS / CAMPANAS.
%
This paper also includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
We thank the TESS Architects (G.~Ricker, R.~Vanderspek, D.~Latham,
S.~Seager, J.~Jenkins) and the many TESS team members for their
efforts to make the mission a continued success.
%
%
%This study was based in part on observations at Cerro Tololo
%Inter-American Observatory at NSF's NOIRLab (NOIRLab Prop{.} ID
%2020A-0146; 2020B-0029 PI: Bouma), which is managed by the
%Association of Universities for Research in Astronomy (AURA) under a
%cooperative agreement with the National Science Foundation.
%
%
Finally, this research has made use of the Keck Observatory Archive (KOA),
which is operated by the W. M. Keck Observatory and the NASA Exoplanet
Science Institute (NExScI), under contract with the National
Aeronautics and Space Administration.  We also thank the Keck
Observatory staff for their support of HIRES and remote observing.  We
recognize the importance that the summit of Maunakea has within the
indigenous Hawaiian community, and are deeply grateful to have the
opportunity to conduct observations from this mountain.
%
% The Digitized Sky Survey was produced at the Space Telescope Science
% Institute under U.S. Government grant NAG W-2166.
% Figure~\ref{fig:scene} is based on photographic data obtained using
% the Oschin Schmidt Telescope on Palomar Mountain.
%

@@

\input{starparams.tex}
\input{model_posterior_table.tex}

@@

The \texttt{v0.5} CDIPS target catalog (Table~\ref{tab:v05}) includes
some important updates from previous versions.  As in
\citet{bouma_cdipsI_2019}, we collected membership information for
young, age-dated, or age-dateable stars from across the literature.
Table~\ref{tab:metadata} gives a list of the sources included, and
some brief summary statistics.

@@

\input{t1_appendix.tex}
\input{t2_appendix.tex}

>>>>

The K2 and TESS missions have now enabled the detection of about ten
close-in planets younger than 100 million years, all smaller than
Jupiter
\citep{Mann_K2_33b_2016,David_et_al_2017,david_four_2019,newton_tess_2019,bouma_cluster_2020,plavchan_planet_2020,rizzuto_tess_2020}.
The Kepler mission however has not yielded any planets with precise
ages below one gigayear \citep{Meibom_et_al_2013}.  The reason is that
during the main Kepler mission (2009--2013), only four open
clusters were known in the Kepler field: NGC\,6866, NGC\,6811,
NGC\,6819, and NGC\,6791, with ages spanning 0.7\,Gyr to 9\,Gyr
\citep{meibom_kepler_2011}.  Since that time, analyses of the
kinematic, photometric, and astrometric Gaia data have expanded our
knowledge of open cluster and moving group memberships \citep[{\it
e.g.},][]{cantatgaudin_gaia_2018,zari_3d_2018,kounkel_untangling_2019,Meingast2021,Kerr2021}.
As part of our Cluster Difference Imaging Photometric Survey (CDIPS,
\citealt{bouma_cdipsI_2019}), we concatenated the available analyses
from the literature, which yielded a list of candidate young and
age-dated stars (see Appendix~\ref{app:targetlist}).

@@

Ultimately, the main advance of this work was a precise measurement of
the age of \pn.   This age measurement was enabled by identifying the
connection of the star to the \cn\ using the Gaia data.  Table~\ref{tab:v06}
should enable similar crossmatches to
already known planets, as well as forthcoming planets from the TESS
mission \citep{ricker_transiting_2015,guerrero_tess_2021}.  An
alternative path toward age-dating more planets could be to identify kinematic associations
around known exoplanet host stars using positions and tangential velocities
from Gaia, and to then verify these associations with ancilliary
rotation periods and spectroscopy \citep[{\it e.g.},][]{tofflemire_tess_2021}.

@@

L.G.B{.} and J.L.C{.} are grateful to T{.}~David for help with the
transit fitting, to R{.}~Kerr for kindly providing us with the
\citet{Kerr2021} membership list prior to its publication.
L.G.B{.}, L.H{.}, and N.N{.} are also grateful to K{.}~Collins for
helping resolve the scheduling conflict that would have otherwise
prevented the MuSCAT3 observations.
% is grateful to G.~Zhou, B{.}~Tofflemire, A{.}~McWilliam,
% E{.}~Newton, M{.}~Kounkel, A{.}~Kraus, L{.}~Hillenbrand, and
% K{.}~Hawkins for the discussions on young stars, rotation, and lithium
% that encouraged this analysis.
%
L.G.B{.} acknowledges support by the TESS GI Program, programs
G011103 and G022117, through NASA grants 80NSSC19K0386 and
80NSSC19K1728.
%
L.G.B{.} was also supported by a Charlotte Elizabeth Procter
Fellowship from Princeton University.
%
The archival Keck/NIRC2 imaging was acquired by programs 2015A/N301N2L
(PI: A.~Kraus) and 2019A/N069 (PI: E.~Petigura).
%
In addition, this paper is based in part on observations made with the
MuSCAT3 instrument, developed by the Astrobiology Center and under
financial supports by JSPS KAKENHI (JP18H05439) and JST PRESTO
(JPMJPR1775), at Faulkes Telescope North on Maui, HI, operated by the
Las Cumbres Observatory.
%
This work is partly supported by JSPS KAKENHI Grant Numbers 22000005,
JP15H02063, JP17H04574, JP18H05439, JP18H05442, JST PRESTO Grant
Number JPMJPR1775, the Astrobiology Center of National Institutes of
Natural Sciences (NINS) (Grant Number AB031010).
%
% ACKNOWLEDGE PFS / CAMPANAS.
%
This paper also includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
We thank the TESS Architects (G.~Ricker, R.~Vanderspek, D.~Latham,
S.~Seager, J.~Jenkins) and the many TESS team members for their
efforts to make the mission a continued success.
%
%
%This study was based in part on observations at Cerro Tololo
%Inter-American Observatory at NSF's NOIRLab (NOIRLab Prop{.} ID
%2020A-0146; 2020B-0029 PI: Bouma), which is managed by the
%Association of Universities for Research in Astronomy (AURA) under a
%cooperative agreement with the National Science Foundation.
%
%
Finally, this research has made use of the Keck Observatory Archive (KOA),
which is operated by the W. M. Keck Observatory and the NASA Exoplanet
Science Institute (NExScI), under contract with the National
Aeronautics and Space Administration.  We also thank the Keck
Observatory staff for their support of HIRES and remote observing.  We
recognize the importance that the summit of Maunakea has within the
indigenous Hawaiian community, and are deeply grateful to have the
opportunity to conduct observations from this mountain.
%
% The Digitized Sky Survey was produced at the Space Telescope Science
% Institute under U.S. Government grant NAG W-2166.
% Figure~\ref{fig:scene} is based on photographic data obtained using
% the Oschin Schmidt Telescope on Palomar Mountain.
%

@@

\input{starparams.tex}
\input{model_posterior_table.tex}
\input{t1_appendix.tex}
\input{t2_appendix.tex}

@@

The \texttt{v0.6} CDIPS target catalog (Table~\ref{tab:v06}) includes
some important updates from previous versions.  As in
\citet{bouma_cdipsI_2019}, we collected membership information for
young, age-dated, or age-dateable stars from across the literature.
Table~\ref{tab:metadata} gives a list of the sources included, and
some brief summary statistics.

@@

### 2021-09-07 paper/t1_appendix.tex
update star table

%% This is the title of the table.
\tablecaption{Young, Age-dated, and Age-dateable Stars Within the
  Nearest Few Kiloparsecs ($\texttt{v0.5}$ of the CDIPS Target
  List).}
\label{tab:v05}

@@

%% All data must appear between the \startdata and \enddata commands
%
% paste from
% /Users/luke/Dropbox/proj/rudolf/results/tables/v05_main_tableheader.tex
% via drivers/write_v05_main_tableheader.py
\startdata
         \texttt{source\_id} &                                         1709456705329541504 &                                             Gaia DR2 source identifier. \\
                 \texttt{ra} &                                                     247.826 &                                         Gaia DR2 right ascension [deg]. \\
                \texttt{dec} &                                                      79.789 &                                             Gaia DR2 declination [deg]. \\
           \texttt{parallax} &                                                      35.345 &                                                Gaia DR2 parallax [mas]. \\
    \texttt{parallax\_error} &                                                       0.028 &                                    Gaia DR2 parallax uncertainty [mas]. \\
               \texttt{pmra} &                                                      94.884 &     Gaia DR2 proper motion $\mu_\alpha \cos \delta$ [mas$\,$yr${^-1}$]. \\
              \texttt{pmdec} &                                                     -86.971 &                 Gaia DR2 proper motion $\mu_\delta$ [mas$\,$yr${^-1}$]. \\
 \texttt{phot\_g\_mean\_mag} &                                                        6.85 &                                                 Gaia DR2 $G$ magnitude. \\
\texttt{phot\_bp\_mean\_mag} &                                                       6.409 &                                     Gaia DR2 $G_\mathrm{BP}$ magnitude. \\
\texttt{phot\_rp\_mean\_mag} &                                                       7.189 &                                     Gaia DR2 $G_\mathrm{RP}$ magnitude. \\
            \texttt{cluster} &                 Uma,IR\_excess,NASAExoArchive\_ps\_20210506 &                                  Comma-separated cluster or group name. \\
                \texttt{age} &                                                nan,nan,9.48 & Comma-separated logarithm (base-10) of reported$^{\rm a}$ age in years. \\
          \texttt{mean\_age} &                                                        9.48 &                          Mean (ignoring NaNs) of $\texttt{age}$ column. \\
      \texttt{reference\_id} &      Ujjwal2020,CottenSong2016,NASAExoArchive\_ps\_20210506 &                          Comma-separted provenance of group membership. \\
 \texttt{reference\_bibcode} & 2020AJ....159..166U,2016ApJS..225...15C,2013PASP..125..989A &                  ADS bibcode corresponding to $\texttt{reference\_id}$. \\
\enddata

@@

%% General table comment marker
\tablecomments{
Table~\ref{tab:v05} is
published in its entirety in a machine-readable format.   This table is a
concatenation of the studies listed in Table~\ref{tab:metadata}.
One entry is
shown for guidance regarding form and content. 
In this particular example, the star has a
cold Jupiter on a 16 year orbit, HD 150706b \citep{2012AA...545A..55B}.
An infrared excess has been reported \citep{CottenSong2016}, and the star was identified by \citet{Ujjwal2020} as a candidate UMa moving group
member ($\approx 400\,{\rm Myr}$; \citealt{mann_tess_2020}).
The star's RV activity and TESS rotation period corroborate its youth.
}
\vspace{-0.5cm}
\end{deluxetable*}

>>>>

%% This is the title of the table.
\tablecaption{Young, Age-dated, and Age-dateable Stars Within the
  Nearest Few Kiloparsecs.}
\label{tab:v06}

@@

%% All data must appear between the \startdata and \enddata commands
%
% paste from
% /Users/luke/Dropbox/proj/rudolf/results/tables/v06_main_tableheader.tex
% via drivers/write_v06_main_tableheader.py
\startdata
          \texttt{source\_id} &                                          1709456705329541504 &                                              Gaia DR2 source identifier. \\
                  \texttt{ra} &                                                      247.826 &                                          Gaia DR2 right ascension [deg]. \\
                 \texttt{dec} &                                                       79.789 &                                              Gaia DR2 declination [deg]. \\
            \texttt{parallax} &                                                       35.345 &                                                 Gaia DR2 parallax [mas]. \\
     \texttt{parallax\_error} &                                                        0.028 &                                     Gaia DR2 parallax uncertainty [mas]. \\
                \texttt{pmra} &                                                       94.884 &      Gaia DR2 proper motion $\mu_\alpha \cos \delta$ [mas$\,$yr$^{-1}$]. \\
               \texttt{pmdec} &                                                      -86.971 &                  Gaia DR2 proper motion $\mu_\delta$ [mas$\,$yr$^{-1}$]. \\
  \texttt{phot\_g\_mean\_mag} &                                                         6.85 &                                                  Gaia DR2 $G$ magnitude. \\
 \texttt{phot\_bp\_mean\_mag} &                                                        6.409 &                                      Gaia DR2 $G_\mathrm{BP}$ magnitude. \\
 \texttt{phot\_rp\_mean\_mag} &                                                        7.189 &                                      Gaia DR2 $G_\mathrm{RP}$ magnitude. \\
             \texttt{cluster} &                  NASAExoArchive\_ps\_20210506,Uma,IR\_excess &                                   Comma-separated cluster or group name. \\
                 \texttt{age} &                                                 9.48,nan,nan &  Comma-separated logarithm (base-10) of reported$^{\rm a}$ age in years. \\
           \texttt{mean\_age} &                                                         9.48 &                           Mean (ignoring NaNs) of $\texttt{age}$ column. \\
       \texttt{reference\_id} &       NASAExoArchive\_ps\_20210506,Ujjwal2020,CottenSong2016 &                           Comma-separted provenance of group membership. \\
  \texttt{reference\_bibcode} &  2013PASP..125..989A,2020AJ....159..166U,2016ApJS..225...15C &                   ADS bibcode corresponding to $\texttt{reference\_id}$. \\
\enddata

@@

%% General table comment marker
\tablecomments{
Table~\ref{tab:v06} is published in its entirety in a machine-readable
format.   This table is a concatenation of the studies listed in
Table~\ref{tab:metadata}.  One entry is shown for guidance regarding
form and content.  In this particular example, the star has a cold
Jupiter on a 16 year orbit, HD 150706b \citep{2012AA...545A..55B}.  An
infrared excess has been reported \citep{CottenSong2016}, and the star
was identified by \citet{Ujjwal2020} as a candidate UMa moving group
member ($\approx 400\,{\rm Myr}$; \citealt{mann_tess_2020}).  The
star's RV activity and TESS rotation period corroborate its youth.
}
\vspace{-0.5cm}
\end{deluxetable*}
